\documentclass[a4paper,11pt]{article}
\usepackage{fullpage}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[normalem]{ulem}
\usepackage[english]{babel}
\usepackage{listings,babel}
\lstset{breaklines=true,basicstyle=\ttfamily}
\usepackage{graphicx}
\usepackage{moreverb}
\usepackage{url}

\title{RC5 receiver core}
\author{S\'ebastien Bourdeauducq}
\date{November 2010}
\begin{document}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\maketitle{}
\section{RC5 receiver core}
This core decodes serial data received from an infrared remote control using the RC5 protocol.

It is a fully synchronous design, and uses a clock divider to generate the baud rate. The system clock frequency must be provided using the \verb!clk_freq! parameter.

It features a minimal CPU interface:
\begin{itemize}
\item The core implements one register on the CSR that contains the received RC5 code. The register is 13-bit wide, and contains the 14-bit RC5 code stripped of its first start bit.
\item An interrupt line is pulsed every time a new RC5 code is received.
\end{itemize}

\section*{Copyright notice}
Copyright \copyright 2007-2010 S\'ebastien Bourdeauducq. \\
Permission is granted to copy, distribute and/or modify this document under the terms of the GNU Free Documentation License, Version 1.3; with no Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts. A copy of the license is included in the LICENSE.FDL file at the root of the Milkymist source distribution.

\end{document}
